Bandit Multiclass Classification

### 2018-08-14 main.tex
Add discussions on one-sided banditron and selectron

\section{Problem Statement}
A multiclass linear classifier is defined by a matrix $W\in \mathbb{R}^{K\times d}$, where $K$ is the number of classes ($K\geq 3$) and $d$ is the dimension of the feature vector. A set of samples $\{(x_t, y_t)\}_{t=1}^T$ where $x_t\in \mathbb{R}^d$ and $y_t\in[K]$ is called \textit{linearly separable} if there exists a $W^*$ such that $ y_t= \argmax_{y\in[K]} (W^*x_t)_y $ for all $t\in[T]$. (Note that $(W^*x_t)_y$ can also be written as $\e_y^\top W^* x_t$)

A set of samples $\{(x_t, y_t)\}_{t=1}^T$ is called \textit{linearly separable with a margin $\gamma$} if there exists a $W^*$ with $\|\e_i^\top W^*\|_2\leq 1$ for all $i$ and $(W^*x_t)_{y_t} \geq \max_{y\neq y_t} (W^*x_t)_{y} + \gamma$ for all $t$. For simplicity, we define the set where $W^*$ lies in as $\calW = \{W\in \mathbb{R}^{K\times d}: \|\e_i^\top W^*\|_2\leq 1\}$.

@@

\begin{algorithm}[H]
\caption{Perceptron}
$W_1=\mathbf{0}$\\
\For{$t=1$ to $T$}{
   Predict $\tilde{y}_t=\argmax_y (W_tx_t)_y$ \\
   Update $W_{t+1}\leftarrow W_t + (\e_{y_t}-\e_{\tilde{y}_t})x_t^\top$. 
}
\end{algorithm}
The analysis is simple. On one hand we have (note $\langle A,B \rangle:=\text{Tr}(A^\top B)$):
\begin{align*}
    \langle W_t, W^* \rangle &= \langle W_{t-1}, W^* \rangle + \langle (\e_{y_t}-\e_{\tilde{y}_t})x_t^\top, W^* \rangle \\
    &= \langle W_{t-1}, W^* \rangle + (\e_{y_t}^\top W^* x_t-\e_{\tilde{y}_t}^\top W^* x_t) \\
    &\geq \langle W_{t-1}, W^* \rangle + \gamma\one[\tilde{y}_t\neq y_t]. 
\end{align*}
By induction, we get $\langle W_{T+1}, W^* \rangle \geq \gamma\sum_{t=1}^T \one[\tilde{y}_t\neq y_t]$.

On the other hand, 
\begin{align*}
    \langle W_{T+1}, W^* \rangle \leq \|W_{T+1}\|_F\|W^*\|_F \leq \sqrt{2K\sum_{t=1}^T \one[\tilde{y}_t \neq y_t] \|x_t\|_2^2}\times 1 \leq \sqrt{2K\sum_{t=1}^T \one[\tilde{y}_t \neq y_t] } 
\end{align*}
Combining the above two inequality, we get $ \sum_{t=1}^T \one[\tilde{y}_t \neq y_t] \leq \frac{4K}{\gamma^2}$.

@@

\section{One-versus-all Perceptron}
Our assumption of linearly separable with a margin is 
\begin{align}
 (W^*x_t)_{y_t} \geq (W^*x_t)_{y} + \gamma, \ \forall y\neq y_t. \label{eqn:weaker}
\end{align}
As discussed in Section~\ref{section: difficulty}, it seems hard to make update when $\tilde{y}_t\neq y_t$ and achieve a constant and polynomial error bound.

For a moment, in this subsection we make the following stronger margin assumption (which we call \textit{one-versus-all separable} with a margin): 
\begin{align}
\begin{cases}
    (W^*x_t)_{y_t} \geq \gamma/2 \\
    (W^*x_t)_y \leq -\gamma/2, \forall y\neq y_t.
\end{cases}
\label{eqn:stronger}
\end{align}
Clearly, one-versus-all separability implies linearly separability, but not the other way around. With one-versus-all separability assumption, we can view the problem as $K$ parallel binary classification problem. The following algorithm achieves constant error bound: 
\begin{algorithm}[H]
\caption{One-versus-all Perceptron}
\label{alg:ova}
Initialize $w_t^{(1)}=\ldots=w_t^{(K)}=\mathbf{0}\in \mathbb{R}^d$\\
\For{$t=1$ to $T$}{
   \If{$\exists y$ such that $w_t^{(y)\top} x_t \geq 0$}{
       Assign $\tilde{y}_t$ to any $y$ with $w_t^{(y)\top}  x_t \geq 0$ \\
       Predict $\tilde{y}_t$ \\
       \lIf{$\tilde{y}_t\neq y_t$}{ update $w_t^{(\tilde{y}_t)} \leftarrow w_t^{(\tilde{y}_t)} - x_t$ } \label{line:update1}
   }
   \Else{
       Pick $\tilde{y}_t$ randomly from $
       \text{unif}[K]$ \label{line:explore}\\
       Predict $\tilde{y}_t$ \\
       \lIf{$\tilde{y}_t = y_t$}{ update $w_t^{(\tilde{y}_t)} \leftarrow w_t^{(\tilde{y}_t)} + x_t$ } \label{line:update2}
   }
}
\end{algorithm}
\textbf{Analysis}. Note that when the algorithm enters Line~\ref{line:update1} or Line~\ref{line:update2}, the binary classifier $w_t^{(\tilde{y}_t)}$ is making an error. Let the number of times the algorithm enters Line~\ref{line:update1} and Line~\ref{line:update2} be $M$ and $N$ respectively. By the error bound of binary perceptron $\mathcal{O}\left(\frac{1}{\gamma^2}\right)$, we have that $M+N=\mathcal{O}\left(\frac{K}{\gamma^2}\right)$. Then note that the number of times the algorithm makes a mistake (i.e. $\tilde{y}_t\neq y_t$) is upper bounded by $M$ plus how many times the algorithm explores in Line~\ref{line:explore}; and when the algorithm explores, with probability $\frac{1}{K}$ it enters Line~\ref{line:update2}. Therefore, the number of mistakes is bounded (in expectation) by $\E[M+KN]=\mathcal{O}\left(\frac{K^2}{\gamma^2}\right)$.

\section{One-versus-all Kernel Perceptron}
With the polynomial kernel defined in \cite{klivans2004learning}, we can transform the weaker assumption \eqref{eqn:weaker} in the original feature space to the stronger one \eqref{eqn:stronger} in a transformed feature space. Indeed, in the original feature space, class $i$ corresponds to an intersection of $K-1$ halfspaces: $ \{x: (w_i^*-w_j^*)^\top x \geq 0, \forall j\neq i\} $, which is the subject of \cite{klivans2004learning} (here, $w_i^{*\top}$ equals to $\e_i^\top W^*$ defined above). Directly using their kernel construction, we can have margin $ \left(\frac{\gamma}{d}\right)^{\Omega(K\log K\log (1/\gamma))} $ in the transformed feature space. Running kernelized version of Algorithm~\ref{alg:ova}, we can get $ K^2\left(\frac{d}{\gamma}\right)^{\mathcal{O}(K\log K\log (1/\gamma))} $ error bound directly.

\subsection{Refinement for \cite{klivans2004learning}'s Fact 1}
\begin{lemma}
   For $i=1,\ldots, \ell$ let $q^{(i)}(x)=\sum_S c^{(i)}_{S} x_S $ be a polynomial over $x_1, \ldots, x_d$ with $\|q^{(i)}\|^2\leq M_i$. Then (1) if $q^{(1)}\ldots q^{(\ell)}$ has degree at most $\deg$, we have $\|q^{(1)}\ldots q^{(\ell)}\|^2 \leq (\deg)^{O(\deg)} \prod_i M_i$, and (2) we have $\|q^{(1)}+\ldots+q^{(\ell)}\|^2\leq \ell(M_1+\ldots+M_\ell)$. 
\end{lemma}
\begin{proof}
For the first bound, we bound the ratio between the following two values:  $ \|q^{(1)}\ldots q^{(\ell)}\|^2 $ and $\|q^{(1)}\|^2\ldots \|q^{(\ell)}\|^2$. 
\begin{align}
    n_1&=\prod_{i=1}^\ell \|q^{(i)}\|^2 
    = \prod_{i=1}^\ell \left(\sum_{S_i} \left(c^{(i)}_{S_i}\right)^2 \right)
    = \sum_{(S_1,\ldots,S_\ell)}  \left(\prod_{i=1}^\ell \left(c^{(i)}_{S_i}\right)^2 \right)
    = \sum_{(S_1,\ldots,S_\ell)}  \left(\prod_{i=1}^\ell c^{(i)}_{S_i}\right)^2 \nonumber\\
    &= \sum_S \sum_{\substack{(S_1,\ldots,S_\ell): \\ S_1\times\ldots S_\ell=S}}  \left(\prod_{i=1}^\ell c^{(i)}_{S_i}\right)^2
\end{align}
\begin{align}
    n_2=\|q^{(1)}\ldots q^{(\ell)}\|^2 = \sum_{S} c_S^2 
    = \sum_S \left( \sum_{\substack{(S_1,\ldots, S_\ell): \\ S_1\times\ldots\times S_\ell=S }} \left(c_{S_1}^{(1)}\ldots c_{S_\ell}^{(\ell)} \right) \right)^2 
    = \sum_S \left( \sum_{\substack{(S_1,\ldots, S_\ell): \\ S_1\times\ldots\times S_\ell=S }} \prod_{i=1}^\ell c_{S_i}^{(i)} \right)^2 
\end{align}
Let $M$ be an upper bound of the number of terms involved in the summation $ \sum_{\substack{(S_1,\ldots, S_\ell): \\ S_1\times\ldots\times S_\ell=S }} $, then by Cauchy-Schwarz's inequality we have $ n_2\leq Mn_1 $.

@@

\subsection{Refinement for \cite{klivans2004learning}'s Theorem 7}
\begin{lemma}
Under the same condition as in \cite{klivans2004learning}'s Theorem 7, the PTF has margin on $X$ is at least $\left( \frac{\rho}{t\log t \log \frac{1}{\rho}} \right)^{O(t\log t\log 1/\rho)}$. 
\end{lemma}
\begin{proof}
Using the same construction, we bound $\|p\|$. First, we have $\|\frac{2w^i\cdot x}{\rho}\|^2\leq \frac{4}{\rho^2}$ and $\|\left(\frac{2w^i\cdot x}{\rho}\right)^j\|^2\leq j^j\left(\frac{4}{\rho^2}\right)^j=\left(\frac{4j}{\rho^2}\right)^j$. $a(x), b(x)$ are polynomials of degree $O(\log t\log \frac{1}{\rho})$ with coefficients of magnitude $\left(\frac{1}{\rho}\right)^{O(\log t\log 1/\rho)}$. Thus 
\begin{align*}
    \|a(2w^i\cdot x/\rho)\|^2 &\leq 
    \left( \frac{\log t \log \frac{1}{\rho}}{\rho} \right)^{O(\log t\log 1/\rho)}  \times  \left(\frac{1}{\rho}\right)^{O(\log t\log 1/\rho)}\\
    &\leq \left( \frac{\log t \log \frac{1}{\rho}}{\rho} \right)^{O(\log t\log 1/\rho)}.
\end{align*}
Same holds for $\|b(2w^i\cdot x/\rho)\|^2$. Finally we have $\|B(x)\|^2 \leq (t)^{O(t\log t\log 1/\rho)} \times \left( \frac{\log t \log \frac{1}{\rho}}{\rho} \right)^{O(t\log t\log 1/\rho)} = \left( \frac{t\log t \log \frac{1}{\rho}}{\rho} \right)^{O(t\log t\log 1/\rho)}$
\end{proof}

>>>>

\section{Problem Statement}
A multiclass linear classifier is defined by a set of $K$ vectors $w^{(1)}, \ldots, w^{(K)}\in \mathbb{R}^d$ (or a compact matrix form $W = [w^{(1)}\ w^{(2)}\  \cdots\  w^{(K)}]^\top \in \mathbb{R}^{K\times d}$), where $K$ is the number of classes ($K\geq 3$) and $d$ is the dimension of the feature vector. A set of samples $\{(x_t, y_t)\}_{t=1}^T$ where $x_t\in \mathbb{R}^d$ and $y_t\in[K]$ is called \textit{linearly separable} if there exists a $w_*^{(1)}, \ldots, w_*^{(K)}$ (or simply $W_*$) such that $ y_t= \argmax_{y\in[K]} (W_{*}x_t)_y $ for all $t\in[T]$.

A set of samples $\{(x_t, y_t)\}_{t=1}^T$ is called \textit{linearly separable with a margin $\gamma$} if there exists $W_*$ with $\|w^{(i)}_*\|_2\leq 1$ for all $i$ and $(W_{*}x_t)_{y_t} \geq \max_{y\neq y_t} (W_*x_t)_y + \gamma$ for all $t$. We define the set $\calW = \{W\in \mathbb{R}^{K\times d}: \|\e_i^\top W^*\|_2\leq 1\}$. Therefore, $W_*\in \calW$.

@@

\begin{algorithm}[H]
\caption{Perceptron}
$W_1=\mathbf{0}$\\
\For{$t=1$ to $T$}{
   Predict $\tilde{y}_t=\argmax_y (W_tx_t)_y$ \\
   Update $W_{t+1}\leftarrow W_t + (\e_{y_t}-\e_{\tilde{y}_t})x_t^\top$ \\
   (same as updating $w^{(y_t)}_{t+1} \leftarrow w^{(y_t)}_t + x_t$ and $w^{(\hat{y}_t)}_{t+1} \leftarrow w^{(\hat{y}_t)}_t - x_t$ when $\hat{y}_t\neq y_t$). 
}
\end{algorithm}
The analysis is simple. On one hand we have (note $\langle A,B \rangle:=\text{Tr}(A^\top B)$):
\begin{align*}
    \langle W_t, W_* \rangle &= \langle W_{t-1}, W_* \rangle + \langle (\e_{y_t}-\e_{\tilde{y}_t})x_t^\top, W_* \rangle \\
    &= \langle W_{t-1}, W_* \rangle + (\e_{y_t}^\top W_* x_t-\e_{\tilde{y}_t}^\top W_* x_t) \\
    &\geq \langle W_{t-1}, W_* \rangle + \gamma\one[\tilde{y}_t\neq y_t]. 
\end{align*}
By induction, we get $\langle W_{T+1}, W_* \rangle \geq \gamma\sum_{t=1}^T \one[\tilde{y}_t\neq y_t]$.

On the other hand, 
\begin{align*}
    \langle W_{T+1}, W_* \rangle \leq \|W_{T+1}\|_F\|W_*\|_F \leq \sqrt{2K\sum_{t=1}^T \one[\tilde{y}_t \neq y_t] \|x_t\|_2^2}\times 1 \leq \sqrt{2K\sum_{t=1}^T \one[\tilde{y}_t \neq y_t] } 
\end{align*}
Combining the above two inequality, we get $ \sum_{t=1}^T \one[\tilde{y}_t \neq y_t] \leq \frac{4K}{\gamma^2}$.

@@

\section{One-versus-all Perceptron}
\label{sec:ova}
Our assumption of linearly separable with a margin is 
\begin{align}
 (W_*x_t)_{y_t} \geq (W_*x_t)_{y} + \gamma, \ \forall y\neq y_t. \label{eqn:weaker}
\end{align}
As discussed in Section~\ref{section: difficulty}, it seems hard to make update when $\tilde{y}_t\neq y_t$ and achieve a constant and polynomial error bound.

For a moment, in this section we make the following stronger margin assumption (which we call \textit{one-versus-all separable} with a margin): 
\begin{align}
\begin{cases}
    (W_*x_t)_{y_t} \geq \gamma/2 \\
    (W_*x_t)_y \leq -\gamma/2, \forall y\neq y_t.
\end{cases}
\label{eqn:stronger}
\end{align}
Clearly, one-versus-all separability implies linearly separability, but not the other way around. With one-versus-all separability assumption, we can view the problem as $K$ parallel binary classification problem. The following algorithm achieves constant error bound: 
\begin{algorithm}[H]
\caption{One-versus-all Perceptron}
\label{alg:ova}
Initialize $w_t^{(1)}=\ldots=w_t^{(K)}=\mathbf{0}\in \mathbb{R}^d$\\
\For{$t=1$ to $T$}{
   \If{$\exists y$ such that $w_t^{(y)\top} x_t \geq 0$}{
       Assign $\tilde{y}_t$ to any $y$ with $w_t^{(y)\top}  x_t \geq 0$ \\
       Predict $\tilde{y}_t$ \\
       \lIf{$\tilde{y}_t\neq y_t$}{ update $w_t^{(\tilde{y}_t)} \leftarrow w_t^{(\tilde{y}_t)} - x_t$ } \label{line:update1}
   }
   \Else{
       Pick $\tilde{y}_t$ randomly from $
       \text{unif}[K]$ \label{line:explore}\\
       Predict $\tilde{y}_t$ \\
       \lIf{$\tilde{y}_t = y_t$}{ update $w_t^{(\tilde{y}_t)} \leftarrow w_t^{(\tilde{y}_t)} + x_t$ } \label{line:update2}
   }
}
\end{algorithm}
\textbf{Analysis}. Note that when the algorithm enters Line~\ref{line:update1} or Line~\ref{line:update2}, the binary classifier $w_t^{(\tilde{y}_t)}$ is making an error. Let the number of times the algorithm enters Line~\ref{line:update1} and Line~\ref{line:update2} be $M$ and $N$ respectively. By the error bound of binary perceptron $\mathcal{O}\left(\frac{1}{\gamma^2}\right)$, we have that $M+N=\mathcal{O}\left(\frac{K}{\gamma^2}\right)$. Then note that the number of times the algorithm makes a mistake (i.e. $\tilde{y}_t\neq y_t$) is upper bounded by $M$ plus how many times the algorithm explores in Line~\ref{line:explore}; and when the algorithm explores, with probability $\frac{1}{K}$ it enters Line~\ref{line:update2}. Therefore, the number of mistakes is bounded (in expectation) by $\E[M+KN]=\mathcal{O}\left(\frac{K^2}{\gamma^2}\right)$.

\section{One-versus-all Kernel Perceptron}
%With the polynomial kernel defined in \cite{klivans2004learning}, we can transform the weaker assumption \eqref{eqn:weaker} in the original feature space to the stronger one \eqref{eqn:stronger} in a transformed feature space. Indeed, in the original feature space, class $i$ corresponds to an intersection of $K-1$ halfspaces: $ \{x: (w_i^*-w_j^*)^\top x \geq 0, \forall j\neq i\} $, which is the subject of \cite{klivans2004learning} (here, $w_i^{*\top}$ equals to $\e_i^\top W^*$ defined above). Directly using their kernel construction, we can have margin $ \left(\frac{\gamma}{d}\right)^{\Omega(K\log K\log (1/\gamma))} $ in the transformed feature space. Running kernelized version of Algorithm~\ref{alg:ova}, we can get $ K^2\left(\frac{d}{\gamma}\right)^{\mathcal{O}(K\log K\log (1/\gamma))} $ error bound directly. 
The idea here is to do a feature transformation such that if the original assumption \eqref{eqn:weaker} is satisfied, then in the transformed feature space, \eqref{eqn:stronger} will be satisfied, then we can run the one-versus-all perceptron given in \ref{sec:ova}. Under the original assumption, each class corresponds to an intersection of $K-1$ halfspaces in the feature spaces (i.e., class $i$ corresponds to $\{x\in\mathbb{R}^d: w_*^{(i)\top}x \geq w_*^{(j)\top}x, \ \forall j\neq i\}$). We want this region to become a halfspace in the transformed space. Fortunately, this is possible when there is a margin, and is exactly what is done in \cite{klivans2004learning}.

\subsection{Refinement for \cite{klivans2004learning}'s Fact 1}
\begin{lemma}
   For $i=1,\ldots, \ell$ let $q^{(i)}(x)=\sum_S c^{(i)}_{S} x_S $ be a polynomial over $x_1, \ldots, x_d$ with $\|q^{(i)}\|^2\leq M_i$. Then (1) if $q^{(1)}\ldots q^{(\ell)}$ has degree at most $\deg$, we have $\|q^{(1)}\ldots q^{(\ell)}\|^2 \leq \ell^{\deg} \prod_i M_i$, and (2) we have $\|q^{(1)}+\ldots+q^{(\ell)}\|^2\leq \ell(M_1+\ldots+M_\ell)$. 
\end{lemma}
\begin{proof}
For the first bound, we bound the ratio between the following two values:  $ \|q^{(1)}\ldots q^{(\ell)}\|^2 $ and $\|q^{(1)}\|^2\ldots \|q^{(\ell)}\|^2$. 
\begin{align}
    n_1&=\prod_{i=1}^\ell \|q^{(i)}\|^2 
    = \prod_{i=1}^\ell \left(\sum_{S_i} \left(c^{(i)}_{S_i}\right)^2 \right)
    = \sum_{(S_1,\ldots,S_\ell)}  \left(\prod_{i=1}^\ell \left(c^{(i)}_{S_i}\right)^2 \right)
    = \sum_{(S_1,\ldots,S_\ell)}  \left(\prod_{i=1}^\ell c^{(i)}_{S_i}\right)^2 \nonumber\\
    &= \sum_S \sum_{\substack{(S_1,\ldots,S_\ell): \\ S_1\times\ldots S_\ell=S}}  \left(\prod_{i=1}^\ell c^{(i)}_{S_i}\right)^2
\end{align}
\begin{align}
    n_2=\|q^{(1)}\ldots q^{(\ell)}\|^2 = \sum_{S} c_S^2 
    = \sum_S \left( \sum_{\substack{(S_1,\ldots, S_\ell): \\ S_1\times\ldots\times S_\ell=S }} \left(c_{S_1}^{(1)}\ldots c_{S_\ell}^{(\ell)} \right) \right)^2 
    = \sum_S \left( \sum_{\substack{(S_1,\ldots, S_\ell): \\ S_1\times\ldots\times S_\ell=S }} \prod_{i=1}^\ell c_{S_i}^{(i)} \right)^2 
\end{align}
Let $M$ be an upper bound of the number of terms involved in the summation $ \sum_{\substack{(S_1,\ldots, S_\ell): \\ S_1\times\ldots\times S_\ell=S }} $, then by Cauchy-Schwarz's inequality we have $ n_2\leq Mn_1 $.

@@

\subsection{Refinement for \cite{klivans2004learning}'s Theorem 7}
\begin{lemma}
Under the same condition as in \cite{klivans2004learning}'s Theorem 7, the PTF has margin on $X$ is at least $\left( \frac{\rho}{t\log t \log \frac{1}{\rho}} \right)^{O(t\log t\log 1/\rho)}$. 
\end{lemma}
\begin{proof}
Using the same construction, we bound $\|p\|$. First, we have $\|\frac{2w^i\cdot x}{\rho}\|^2\leq \frac{4}{\rho^2}$ and $\|\left(\frac{2w^i\cdot x}{\rho}\right)^j\|^2\leq j^j\left(\frac{4}{\rho^2}\right)^j=\left(\frac{4j}{\rho^2}\right)^j$. $a(x), b(x)$ are polynomials of degree $O(\log t\log \frac{1}{\rho})$ with coefficients of magnitude $\left(\frac{1}{\rho}\right)^{O(\log t\log 1/\rho)}$. Thus 
\begin{align*}
    \|a(2w^i\cdot x/\rho)\|^2 &\leq 
    \left( \frac{\log t \log \frac{1}{\rho}}{\rho} \right)^{O(\log t\log 1/\rho)}  \times  \left(\frac{1}{\rho}\right)^{O(\log t\log 1/\rho)}\\
    &\leq \left( \frac{\log t \log \frac{1}{\rho}}{\rho} \right)^{O(\log t\log 1/\rho)}.
\end{align*}
Same holds for $\|b(2w^i\cdot x/\rho)\|^2$. Finally we have $\|B(x)\|^2 \leq (t)^{O(t\log t\log 1/\rho)} \times \left( \frac{\log t \log \frac{1}{\rho}}{\rho} \right)^{O(t\log t\log 1/\rho)} = \left( \frac{t\log t \log \frac{1}{\rho}}{\rho} \right)^{O(t\log t\log 1/\rho)}$

%Another refinement is that we can show $|B(x)|\geq $

%\subsection{Possibilities of using other kernels}

\section{A Recipe of Designing Efficient Algorithms}
We have seen that if the data is ``one-versus-all'' separable with a margin, then constant error is achievable. To have constant error, a key is that every time an error occurs, the learner should make significant progress.

Besides one-versus-all separability, are there other kinds of conditions under which we can easily make progress when we make a mistake? If such condition exists, then we have the following recipe of designing algorithms that have constant error bound: 
\begin{itemize}
    \item Design an algorithm that has constant mistake bound under some assumption. 
    \item Design feature transformation such that our original linear separability satisfies the above assumption in the transformed space. 
\end{itemize}
Our kernel perceptron exactly follows the above procedure. Following is another example.

\subsection{One-sided banditron}
Consider the following algorithm. \\
\begin{algorithm}[H]
\caption{Ond-sided banditron}
\label{alg:one-sided}
\textbf{Define}: $\Omega\subseteq \mathbb{R}^{K\times d}$ is a convex set, and we assume $W^* \in \Omega_t$ for all $t$. Learning rate $\eta$. \\
\textbf{Initialize}: $W_1=\mathbf{0}$. \\
\For{$t=1, \ldots, T$}{
   Receive $x_t\in \mathbb{R}^d$. \\
   Predict $\hat{y}_t = \argmax_{y\in[K]} (W_tx_t)_y$. \\
   \If{$\hat{y}_t \neq y_t$}{
       $W_{t+1}'\leftarrow W_t - \eta\e_{\hat{y}_t} x_t^\top$.\\
       $W_{t+1}=\Pi_{\Omega_{t+1}}(W_{t+1}') \triangleq \argmin_{W\in\Omega_{t+1}}\|W-W_{t+1}'\|_F$. 
   }
}
\end{algorithm}
\ \\
The above algorithm is similar to banditron but only performs \textit{one-sided} update and never explicitly explores to get label. Below we check under what condition can this algorithm have constant error bound.

Following standard analysis for online gradient descent, we have when the learner makes a mistake:
\begin{align}
    \|W_{t+1}-W^*\|_F^2 &\leq \|W_{t+1}'-W^*\|_F^2 = \|W_t - W^* - \eta \e_{\hat{y}_t}x_t^\top \|_F^2 \nonumber \\
    &= \|W_t-W^*\|_F^2 - 2\eta ((W_tx_t)_{\hat{y}_t} - (W^*x_t)_{\hat{y}_t}) + \eta^2\|x_t\|_2^2 
    \label{eqn:update_rule}
    %&\leq \|W_t-W^*\|_F^2 - 2\eta ((W_tx_t)_{\hat{y}_t} - (W^*x_t)_{\hat{y}_t}) + \eta^2. 
\end{align}

We see that if the condition 
\begin{align}
    (W_tx_t)_{\hat{y}_t} - (W^*x_t)_{\hat{y}_t} \geq \Delta  \label{eqn:one-sided condition}
\end{align}
holds whenever $\hat{y}_t\neq y_t$ and we pick $\eta = \Delta/\sup_\tau\|x_\tau\|_2^2$, then we always have
\begin{align*}
    \|W_{t+1}-W^*\|_F^2 \leq \|W_{t}-W^*\|_F^2 -\frac{2\Delta^2}{\sup_\tau\|x_\tau\|_2^2} + \frac{\Delta^2}{\sup_\tau\|x_\tau\|_2^2} = \|W_{t}-W^*\|_F^2 - \frac{\Delta^2}{\sup_\tau\|x_\tau\|_2^2} 
\end{align*}
whenever $\hat{y}_t\neq y_t$. Then we have the error bound 
\begin{align*}
    \sum_{t=1}^T \one[\hat{y}_t\neq y_t] \leq \frac{\|W^*\|_F^2 \sup_{\tau}\|x_\tau\|_2^2}{\Delta^2}   
\end{align*}
because $\|W_t - W^*\|_F^2$ is always non-negative and $\|W_1-W^*\|_F^2=\|W^*\|_F^2$. \\

The following analysis leads to the same bound but is more standard. Let $M_t=\one[\hat{y}_t\neq y_t]$. By \eqref{eqn:update_rule} and \eqref{eqn:one-sided condition} we have
\begin{align*}
    \Delta \sum_{t=1}^T M_t  \leq \sum_{t=1}^T M_t\left((W_tx_t)_{\hat{y}_t} - (W^*x_t)_{\hat{y}_t}\right) \leq \frac{\|W^*\|_F^2}{2\eta} + \frac{\eta}{2}\sup_\tau\|x_\tau\|_2^2\sum_{t=1}^T M_t
\end{align*}
Picking appropriate $\eta$ and solving for $\sum_{t=1}^T M_t$ we get the same error bound.

A candidate set of assumptions that leads to \eqref{eqn:one-sided condition} is the \textit{sum-to-zero} assumption: 
\begin{itemize}
    \item $\sum_{i=1}^K \e_i^\top W^* = 0$ (so we pick $\Omega_t = \{W\in \mathbb{R}^{K\times d}: \|\e_i^\top W\|_2\leq 1, \sum_{i=1}^K \e_i^\top W=0\}$) \\
    \item $ (W^*x_t)_y \leq -\Delta $ for all $y\neq y_t$. 
\end{itemize}
They are similar to the assumptions made in \cite{2018arXiv180610745F}. We see that the above two conditions indeed lead to~\eqref{eqn:one-sided condition}: $(W_tx_t)_{\hat{y}_t} = \max_y (W_tx_t)_y \geq \frac{1}{K}\sum_{i=1}^K (W_tx_t)_i =0 $ (recall that we always project $W_t$ back to $\Omega_t$), and when $\hat{y}_t\neq y_t$ we have $(W^*x_t)_{\hat{y}_t}\leq -\Delta$. Unfortunately, this set of assumptions is even stronger than the one-versus-all separable assumption.

Although the one-versus-all assumption is similar, there we do not perform projection for $W_t$, so it is possible that $\max_y (W_tx_t)_y < 0$. It that case, what we do is uniformly sample a label.\\

\textbf{Questions}:
\begin{itemize}
\item Can we avoid the projection step for one-sided banditron (just like perceptron)?
\item Are there other natural assumptions that guarantee \eqref{eqn:one-sided condition}? 
\end{itemize}

\subsection{Algorithms that can update without true labels}
Algorithm~\ref{alg:one-sided} can update without the information of $y_t$. One can think of Algorithm~\ref{alg:one-sided} as a gradient algorithm  on the loss function $\ell_t(W) = \one[\hat{y}_t\neq y_t](Wx_t)_{\hat{y}_t}$. As a comparison, one can think of the surrogate loss function in \cite{2018arXiv180610745F} as $\ell_t(W)=\sum_{y=1}^K \one[y\neq y_t][\Delta+(Wx_t)_y]_+ $ (some scaling and notations are changed for better comparison), whose inverse propensity weighting estimator becomes $\tilde{\ell}_t(W)= \frac{1}{p_{t,\hat{y}_t}}\one[\hat{y}_t\neq y_t][\Delta+(Wx_t)_{
\hat{y}_t}]_+$. One can see that they are closely related, except that in~\cite{2018arXiv180610745F}
when $(Wx_t)_{\hat{y}_t}\leq -\Delta$, the loss for $W$ is set to zero (which makes no difference under the sum-to-zero assumption).

\section{Logarithmic Regret for Strongly Convex Losses}
\subsection{Gradient descent}
In some previous works (we will visit them in the next subsection), it is assumed that the labels are generated with the following process: 
\begin{align}
    p(y|x) = \nabla_y\Phi(W^*x), 
\end{align}
which some $\Phi: \mathbb{R}^K \rightarrow \mathbb{R}$.
The margin assumption is defined as 
\begin{align}
    p(y_t|x_t) - \max_{y\neq y_t}p(y|x_t) \geq \Delta, \forall t.  \label{eqn:new margin}
\end{align}
This margin assumption is a special case of the Tsybakov condition investigated in ~\cite{agarwal2013selective}. 
Defining $\hat{y}_t = \argmax_y \nabla\Phi(W_tx_t)$ and $y^*_t = \argmax_y \nabla\Phi(W^*x_t) $ The regret is defined as
\begin{align}
    \Reg_T &= \sum_{t=1}^T \E\left[ \one[\hat{y}_t\neq y_t] - \one[y_t^*\neq y_t]\right] \nonumber \\
    &= \sum_{t=1}^T \sum_{i=1}^K \left(\one[\hat{y}_t\neq i] - \one[y^*_t\neq i]\right)\nabla_i\Phi(W^*x_t). \label{eqn:new regret}
\end{align}

Below we argue that if $\Phi$ is strongly convex and smooth, then the following algorithm (exactly the algorithm in \cite{agarwal2013selective}) achieves regret logarithmic in $T$.

\begin{assumption}
\begin{align*}
     \Phi(v) + \inner{\nb\Phi(v),u-v} + \frac{\gamma_\ell}{2}\|u-v\|_2^2 \leq \Phi(u)-\Phi(v)\leq \Phi(v) + \inner{\nb\Phi(v),u-v} + \frac{\gamma_u}{2}\|u-v\|_2^2. 
\end{align*}
That is, $\Phi$ is $\gamma_\ell$-strongly convex and $\gamma_u$-smooth. 
\end{assumption}

\begin{algorithm}[H]
\caption{Selectron \cite{agarwal2013selective}}
\textbf{Define}: $\epsilon\triangleq \Delta/(2\gamma_u C')$, where $C'$ is defined in Lemma~\ref{lem:decreasing error} $\eta\triangleq \frac{\sup_{W\in\Omega}\|W\|_F^2}{\gamma_\ell}$, $\Omega\triangleq$ a convex set where $W^*$ lies in.  \\
\textbf{Definition}: $\ell_t(W)=\Phi(Wx_t) - (Wx_t)_{y_t}$.\\
\textbf{Initialization}: $W_1=0, M_1=I$.\\ 
\For{$t=1, \ldots, T$}{
   Observe $x_t$. \\
   \If{$ \| x_t \|_{M_t^{-1}} \geq \epsilon$}{
       Draw $\tilde{y}_t \sim \text{unif}([K])$. 
   }
   \Else{
       Draw $\tilde{y}_t = \hat{y}_t \triangleq \argmax_{y\in[K]}(W_tx_t)_y$. 
   }
   \If{ $\tilde{y}_t=y_t$ }{
       $Z_t\leftarrow 1$, \\
       $M_{t+1}\leftarrow M_t + Z_tx_tx_t^\top$,\\
       $W_{t+1} = \argmin_{W\in\Omega}\left\{ \sum_{s=1}^t Z_s \ell_s(W) + \frac{1}{2\eta}\|W\|_F^2 \right\}$. 
   }
   \Else{
       $Z_t\leftarrow 0$,  \\
       $M_{t+1}\leftarrow M_t$,\\
       $W_{t+1}\leftarrow W_t$.
   }

}
\end{algorithm}

\begin{theorem}
\label{lemma:new main}
The selective sampling algorithm achieves
\begin{align*}
    \Reg_T\leq \tilde{O}\left( \frac{\gamma_u^2}{\gamma_\ell^2}\frac{K^2d}{\Delta^2} \right). 
\end{align*}
\end{theorem}

The proof of this theorem is decomposed to the following three lemmas.

\begin{lemma}[Proposition 1 of \cite{agarwal2013selective}]
\label{lem:decreasing error}
With probability at least $1-\delta$,
\begin{align*}
\|W_t-W^*\|_{M_t} \leq C' \triangleq \frac{2C\sqrt{dK}}{\gamma_\ell} + \sqrt{\frac{2}{\eta \gamma_\ell}}\sup_{W\in \Omega}\|W\|_F, 
\end{align*}
where $C$ hides some $\log t$ and $\log(1/\delta)$ terms (defined in the proof). 
\end{lemma}
